Pre-Main-Sequence M-Dwarfs With Complex Quasiperiodic Variability From Four Years of TESS

### 2023-06-16 paper/ms.tex
update ms

\paragraph{LP 12-502}
While many CPVs had multiple sectors of data, an astounding dataset
was available for LP 12-502 (TIC 402980664; $d$=21\,pc, $J$=9.4,
$T$=11.1).  Figure~\ref{fig:lp} shows all available data from Sectors
18, 19, 25, 26, 53, 58, and 59; the star was observed at 120-second
cadence for every sector during which it was on silicon.

For the first 64 cycles, the star shows a pattern reminiscent of a
(bizarre) eclipsing binary, with four obvious dips, that we dub
dips $\{ 1, 2, 3, 4 \}$ at phases $\{ -0.28, -0.08, 0, 0.25 \}$
respectively.  Over cycles 0 to 64, the depth of dips 1 and 3 remain
roughly fixed.  However dip 4 {\it decreases} in depth by about 2\%,
while dip 2 {\it increases} in depth, by about the same amount.

{\bf todo: discussion needs auto-measured phases at fixed t0...}
There is then a 6-month (184 cycle) gap out to Cycles 248-315, which
show an intricately structured dip complex.
Between Cycles 248-264, there are at least six dips, at phases $\{
  -0.47, -0.28, -0.08,  0.  ,  0.06,  0.17 $\}.
There is ``state switch'', roughly at BTJD 1993.5 - cycle 262.
The lower black points in cycles 248--264 shift upward during cycle
262, never to return.
A large flare occurred a few cycles earlier, at BTJD 1989.2 (which may
or may not be related).
Regardless, the depths of the latter three-dip complex remain roughly
the same after the state-switch;
the first obvious dip in the sequence disappears.
Ah, and the flux changes by 3.5\% over about 22 minutes (0.02\,$P$),
which is the steepest slope exhibited anywhere in this object's
remarkable dataset.
During the beginning of Sector~26 (Cycles 283-298), there are seven
clear dips per cycle.
The the second orbit of this sector,
there is another state-switch at BTJD 2030.7, immediately during a
flare.   (And therefore likely caused by the flare!)

@@

\begin{figure*}[!t]
	\begin{center}
    \centering
    \includegraphics[width=0.9\textwidth]{TIC_402980664_P18.5604_2min_phase_timegroups_mosaic.pdf}
		\vspace{-0.45cm}
		\caption{
			{\it Bottom:}
      Evolution of LP 12-502 ($P$=18.5604\,h) at fixed period and
      epoch over three years; small text denotes cycle number.  
      There are 200 binned black points per cycle.
      The TESS pointing law dictates time gaps; larger gaps tend to
      yield larger shape changes.
      The dips usually evolve over tens to hundreds of cycles.
      However cycles 1233-1264 show a dip that ``switched'' from a depth
      and duration of 3\% and 3\,hr to 0.3\% and 1\,hr over less than
      one cycle.
		}
		\label{fig:lp}
	\end{center}
\end{figure*}

@@

\acknowledgments
This work was supported by the 
Heising-Simons 51~Pegasi~b Fellowship (LGB)

>>>>

\paragraph{LP 12-502}
While many CPVs had multiple sectors of data, an astounding dataset
was available for LP 12-502 (TIC 402980664; $d$=21\,pc, $J$=9.4,
$T$=11.1).  Figure~\ref{fig:lp} shows all available data from Sectors
18, 19, 25, 26, 53, 58, and 59; the star was observed at 120-second
cadence whenever it was observable by TESS.
Measuring the PDM peak period over each sector independently yielded
an average $\langle P \rangle = 18.5560$\,hr, with a difference
between the maximum and minimum sector-level period of about one minute.
Based on this range, the period is therefore stable to at least one
minute ($\pm$0.017\,hr) over the three year baseline.
However, in detail, a period shift of $\pm$1\,minute yields major
phase drifts over the 3 year baseline; that time interval
corresponds to roughly 1/1000$^{\rm th}$ of a period; we have
observed 1500 cycles.
The achievable period precision, $\sigma_{\rm P}$, can be estimated as
\begin{equation}
  \sigma_{\rm P} = \frac { \sigma_{\rm phi} P } { N_{\rm baseline} },
\end{equation}
for $N_{\rm baseline}$ the number of cycles in the observed baseline and
$\sigma_{\rm phi}$ the phase precision with which any one feature (e.g.~a
dip) can be tracked.
Assuming $\sigma_{\rm phi} \approx 0.01$ yields an expected period precision
$\sigma_{\rm P} \approx 0.45 seconds \approx 1.2\times 10^{-4}\,hr$.
By visually fine-tuning over a grid in period, we eventually 
found $P=18.5611 \pm 0.0001$\,hr, which seems to track certain
features in the LP 12-502 light curve over its entire baseline.

What features are observed?
For the first 64 cycles, the star shows a pattern reminiscent of a
(bizarre) eclipsing binary, with four obvious dips, that we dub
dips $\{ 1, 2, 3, 4 \}$ at phases $\{ -0.28, -0.08, 0, 0.25 \}$
respectively.  Over cycles 0 to 64, the depth of dips 1 and 3 remain
roughly fixed.  However dip 4 {\it decreases} in depth by about 2\%,
while dip 2 {\it increases} in depth, by about the same amount.

{\bf todo: discussion needs auto-measured phases at fixed t0...}
There is then a 6-month (184 cycle) gap out to Cycles 248-315, which
show an intricately structured dip complex.
Between Cycles 248-264, there are at least six dips, at phases $\{
  -0.47, -0.28, -0.08,  0.  ,  0.06,  0.17 $\}.
There is ``state switch'', roughly at BTJD 1993.5 - cycle 262.
The lower black points in cycles 248--264 shift upward during cycle
262, never to return.
A large flare occurred a few cycles earlier, at BTJD 1989.2 (which may
or may not be related).
Regardless, the depths of the latter three-dip complex remain roughly
the same after the state-switch;
the first obvious dip in the sequence disappears.
Ah, and the deepest dip shows a flux exercusion of 3.5\% over about 22 minutes (0.02\,$P$),
which is the steepest slope exhibited anywhere in this object's
remarkable dataset.
During Sector~26 (Cycles 283-315), there are seven clear dips per
cycle, and perhaps even eight, if one counts five dips in the largest
dip complex.
During the second orbit of this sector,
there is another state-switch at BTJD 2030.7, immediately during a
flare.   (And therefore likely caused by the flare!)

@@

\begin{figure*}[!t]
	\begin{center}
    \centering
    \includegraphics[width=0.9\textwidth]{TIC_402980664_P18.5604_2min_phase_timegroups_mosaic.pdf}
		\vspace{-0.45cm}
		\caption{
			{\it Bottom:}
      Evolution of LP 12-502 ($P$=18.6\,h) at fixed period and
      epoch over three years; small text denotes cycle number.  
      There are 200 binned black points per cycle.
      The TESS pointing law dictates time gaps; larger gaps tend to
      yield larger shape changes.
      The dips usually evolve over tens to hundreds of cycles.
      However cycles 1233-1264 show a dip that ``switched'' from a depth
      and duration of 3\% and 3\,hr to 0.3\% and 1\,hr over less than
      one cycle.
		}
		\label{fig:lp}
	\end{center}
\end{figure*}

@@

\acknowledgments
This work was supported by the 
Heising-Simons 51~Pegasi~b Fellowship (LGB)
TIC 402980664 was observed at 120-second cadence thanks to the TESS Guest
Investigator programs G022252 (PI: J.~Schlieder; Sectors 18, 19, 25,
26) and G04168 (PI: R.~Jayaraman; Sector 53).
